---
title: "Project 3, Aim 2: Data Processing & Analysis"
author: "WB"
date: "2026-07-17"
output:
  html_document:
    theme: cosmo
    highlight: tango
    toc: true
    toc_float: true
    toc_depth: 3
    code_folding: hide
    df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
rm(list = ls())
library(memisc)
library(haven)
library(epiDisplay)
library(forcats)
library(dplyr)
library(ggplot2)
library(interactions)
library(tidyr)
library(tidyverse)
library(stringr)
library(readxl)
library(table1)
library(knitr)
library(kableExtra)
library(purrr)
library(shiny)
library(reactable)
library(gridExtra)
library(ggbreak)
library(gtsummary)
library(flextable)
library(officer)
library(ipw)
library(survey)
library(pollster)
library(reshape2)
library(multilevel)
library(psych)
library(janitor)
library(rstatix)

# Small helper used throughout to render a data frame as a clean, styled table
pretty_table <- function(df, caption = NULL) {
  df %>%
    kable(caption = caption, digits = 2) %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                  full_width = FALSE, position = "left")
}
```

```{r working directory and calling our file, include=FALSE}
setwd("/Users/will/Documents/UCSD/Project 3/Aim 2/Card Sort Data/")
df_raw  <- read.csv("BaselineSurveyExport_071726.csv")

df  <- as_tibble(df_raw)
```

# Overview

This report summarizes cleaning, scale scoring, reliability, descriptive
statistics, and clinic-level comparisons for the Aim 2 baseline survey
(Implementation Climate, Team Psychological Safety, and Collaboration &
Satisfaction About Care Decisions).

# Initial Cleaning & Renaming

```{r cleaning}
# Remove exported qualtrics labels first two rows
df <- df_raw %>%
  slice(-(1:2))

# Filter if unfinished
df <- df %>%
  filter(Finished == 1)

# Rename other vars
df <- df %>%
  rename(ID = `Last.4.Dig.Phone..`,
         position = `Job.Title`)

# Drop Qualtrics preview/test responses if present
if ("DistributionChannel" %in% names(df)) {
  df <- df %>% filter(DistributionChannel != "preview")
}

df <- df %>%
  mutate(
    Gender = as.factor(Gender),
    Race = as.factor(Race),
    Ethnicity = as.factor(Ethnicity),
    position = as.factor(position),
    Clinic = as.factor(Clinic),
    Experience = suppressWarnings(as.numeric(Experience))
  )

# Clean the clinic names
df <- df %>%
  mutate(
    Clinic_raw = as.character(Clinic),
    Clinic = case_when(
      is.na(Clinic_raw)           ~ NA_character_,
      str_detect(Clinic_raw, ",") ~ "Multiple Clinics",
      Clinic_raw == "1"           ~ "Adolescent Young Adult Medicine",
      Clinic_raw == "2"           ~ "Endocrinology",
      Clinic_raw == "3"           ~ "Gastroenterology",
      Clinic_raw == "4"           ~ "Orthopedics",
      TRUE                        ~ NA_character_
    ),
    Clinic = factor(Clinic, levels = c("Adolescent Young Adult Medicine",
                                        "Endocrinology",
                                        "Gastroenterology",
                                        "Orthopedics",
                                        "Multiple Clinics"))
  )

df <- df %>%
  mutate(
    Gender = factor(Gender,
                     levels = c(1, 2, 7),
                     labels = c("Woman", "Man", "Prefer not to answer")),
    
    Race = factor(Race,
                   levels = c(1, 2, 3, 4, 5, 6, 7, 8),
                   labels = c("White", "Black or African American", 
                              "American Indian or Native Alaskan",
                              "Asian", "Native Hawaiian or Pacific Islander",
                              "More than one", "Other", "Prefer not to answer")),
    
    Ethnicity = factor(Ethnicity,
                        levels = c(1, 2, 3),
                        labels = c("Hispanic", "Non-Hispanic", "Prefer not to answer"))
  )
```

# Scale Scoring

## Implementation Climate (IC)

Scale: 1 (Completely disagree) to 5 (Completely agree). No reverse-coded items.

```{r ic scoring}
ic_items <- paste0("IC_", 1:12)
stopifnot(all(ic_items %in% names(df)))

df <- df %>%
  mutate(across(all_of(ic_items), as.numeric)) %>%
  mutate(
    IC_Focus_Self         = rowMeans(across(c(IC_1, IC_2)), na.rm = TRUE),
    IC_Focus_Clinic        = rowMeans(across(c(IC_7, IC_8)), na.rm = TRUE),
    IC_Support_Self       = rowMeans(across(c(IC_3, IC_4)), na.rm = TRUE),
    IC_Support_Clinic      = rowMeans(across(c(IC_9, IC_10)), na.rm = TRUE),
    IC_Recognition_Self   = rowMeans(across(c(IC_5, IC_6)), na.rm = TRUE),
    IC_Recognition_Clinic  = rowMeans(across(c(IC_11, IC_12)), na.rm = TRUE),
    IC_Total = rowMeans(across(all_of(ic_items)), na.rm = TRUE)
  )
```

## Team Psychological Safety (EPSC, Edmondson)

Scale: 1–7. Items 1, 3, and 5 are reverse-scored.

```{r epsc scoring}
epsc_items <- paste0("EPSC_", 1:7)
stopifnot(all(epsc_items %in% names(df)))

df <- df %>%
  mutate(across(all_of(epsc_items), as.numeric)) %>%
  mutate(
    EPSC_1_r = 8 - EPSC_1,
    EPSC_3_r = 8 - EPSC_3,
    EPSC_5_r = 8 - EPSC_5,
    EPSC_Total = rowMeans(
      across(c(EPSC_1_r, EPSC_2, EPSC_3_r, EPSC_4, EPSC_5_r, EPSC_6, EPSC_7)),
      na.rm = TRUE
    )
  )
```

## Collaboration & Satisfaction About Care Decisions (CSACD)

- **CSACD.I** — Collaboration (6 items, 1–7 scale) → averaged
- **CSACD.II** — Satisfaction with decision (1 item) → used as-is
- **CSACD.III** — Decision-making pattern (2 items, 1–7 scale) → averaged

```{r csacd scoring}
csacd1_items <- paste0("CSACD.I_", 1:6)
stopifnot(all(csacd1_items %in% names(df)))

likert_cols <- c(paste0("IC_", 1:12), paste0("EPSC_", 1:7),
                  paste0("CSACD.I_", 1:6), "CSACD.II_1", "CSACD.III_1", "CSACD.III_2")

df <- df %>%
  mutate(across(all_of(likert_cols), as.numeric)) %>%
  mutate(
    CSACD_Collaboration = rowMeans(across(all_of(csacd1_items)), na.rm = TRUE),
    CSACD_Satisfaction   = CSACD.II_1,
    CSACD_DecisionMaking = rowMeans(across(c(CSACD.III_1, CSACD.III_2)), na.rm = TRUE)
  )
```

# Sample

```{r demographics}
clinic_table <- df %>% tabyl(Clinic) %>%
  mutate(percent = round(percent * 100, 1))
pretty_table(clinic_table, caption = "Respondents by Clinic")


#table 1
table1::table1(~ Gender + Race + Ethnicity + Experience + 
                 IC_Total + EPSC_Total + CSACD_Collaboration | Clinic, 
               data = df)
```



# Reliability (Cronbach's Alpha)

Rather than printing each subscale separately, results are collected into one
tidy, sortable table.

```{r reliability}
alpha_row <- function(data, items, label) {
  a <- suppressWarnings(psych::alpha(data[, items], check.keys = TRUE))
  tibble(
    Scale = label,
    `# Items` = length(items),
    `Raw Alpha` = round(a$total$raw_alpha, 2),
    `Std. Alpha` = round(a$total$std.alpha, 2),
    `Mean` = round(a$total$mean, 2),
    `SD` = round(a$total$sd, 2)
  )
}

reliability_table <- bind_rows(
  alpha_row(df, ic_items, "IC — All 12 items"),
  alpha_row(df, c("IC_1","IC_2","IC_7","IC_8"), "IC — Focus subscale"),
  alpha_row(df, c("IC_3","IC_4","IC_9","IC_10"), "IC — Support subscale"),
  alpha_row(df, c("IC_5","IC_6","IC_11","IC_12"), "IC — Recognition subscale"),
  alpha_row(df, c("EPSC_1_r","EPSC_2","EPSC_3_r","EPSC_4","EPSC_5_r","EPSC_6","EPSC_7"), "EPSC — All 7 items (reverse-coded)"),
  alpha_row(df, csacd1_items, "CSACD.I — Collaboration")
)

pretty_table(reliability_table, caption = "Scale Reliability Summary")
```

# Descriptive Statistics

```{r descriptives}
score_vars <- c("IC_Total", "IC_Focus_Self", "IC_Focus_Clinic",
                 "IC_Support_Self", "IC_Support_Clinic",
                 "IC_Recognition_Self", "IC_Recognition_Clinic",
                 "EPSC_Total", "CSACD_Collaboration", "CSACD_Satisfaction",
                 "CSACD_DecisionMaking")

desc_table <- df %>%
  select(all_of(score_vars)) %>%
  psych::describe() %>%
  as.data.frame() %>%
  rownames_to_column("Variable") %>%
  select(Variable, n, mean, sd, min, max, skew, kurtosis) %>%
  mutate(across(where(is.numeric), ~ round(.x, 2)))

pretty_table(desc_table, caption = "Descriptive Statistics — Composite Scores")
```

# Clinic Comparisons

```{r comparisons}
means_by_clinic <- df %>%
  filter(!is.na(Clinic)) %>%
  group_by(Clinic) %>%
  summarise(across(all_of(score_vars),
                    list(mean = ~mean(.x, na.rm = TRUE),
                         sd = ~sd(.x, na.rm = TRUE)),
                    .names = "{.col}_{.fn}")) %>%
  mutate(across(where(is.numeric), ~ round(.x, 2)))

pretty_table(means_by_clinic, caption = "Mean (SD) of Each Score by Clinic") %>%
  scroll_box(width = "100%")

anova_by_clinic <- map_dfr(score_vars, function(var) {
  res <- df %>%
    filter(!is.na(Clinic)) %>%
    anova_test(as.formula(paste(var, "~ Clinic")))
  res <- as.data.frame(res)
  class(res) <- "data.frame"
  cbind(Variable = var, res)
})

anova_by_clinic <- anova_by_clinic %>%
  mutate(across(where(is.numeric), ~ round(.x, 2)))

pretty_table(anova_by_clinic, caption = "One-Way ANOVA: Each Score by Clinic")
```


##Remember to use this is to update code locally, which will update the files 
cd ~/Documents/UCSD/project3-aim2-report
git pull origin main
git add Descriptivestats.Rmd
git commit -m "Update analysis"
git push

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
